### 2025-01-23 scripts/compare-hard-shell-with-accessory.Rmd
add first results files

# Compare hard shell genes vs. accessory

```{r}
library(ggplot2)
```

```{r}
# Pangenome statistics for PTUs
ptu = read.csv('../data/PTU-pangenome-stats.csv')

# Read in RMES comparisons
df_4 = read.csv('~/Downloads/trieste/test_k4.csv')
df_6 = read.csv('~/Downloads/trieste/test_k6.csv')
df_4_filter = df_4[df_4$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]
df_6_filter = df_6[df_6$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]
```

Do the statistical test

```{r}
wilcox.test(df_6_filter$discrepancy_selected, df_6_filter$discrepancy_others, paired = TRUE)

```
```{r}
ggplot(df_4_filter, aes(x="k=4", wilcoxon_stat, colour=p_value<0.05))+
  geom_point()
```