### 2025-04-03 scripts/Paper-analysis.Rmd
updating analysis

```{r}
library(ggplot2)
library(knitr)
library(xtable)
library(dplyr)
library(tidyr)
```

## Compare hard shell genes vs. accessory

### Short palindromes (k=4, 6)

```{r}
# Pangenome statistics for PTUs
ptu = read.csv('../data/PTU-pangenome-stats.csv')

# Read in RMES comparisons
df_4 = read.csv('~/Downloads/trieste/test_k4.csv')
df_6 = read.csv('~/Downloads/trieste/test_k6.csv')
df_4_filter = df_4[df_4$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]
df_6_filter = df_6[df_6$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]
```

Do the statistical test

```{r}
wilcox.test(df_4_filter$discrepancy_selected, df_4_filter$discrepancy_others, paired = TRUE)
wilcox.test(df_6_filter$discrepancy_selected, df_6_filter$discrepancy_others, paired = TRUE)

```

### All RM targets (not included in main text)

```{r }
df_4 = read.csv('~/Downloads/trieste/test_4mer_targets.csv')
df_4_filter = df_4[df_4$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]
df_5 = read.csv('~/Downloads/trieste/test_5mer_targets.csv')
df_5_filter = df_5[df_5$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]
df_6 = read.csv('~/Downloads/trieste/test_6mer_targets.csv')
df_6_filter = df_6[df_6$PTU %in% ptu$PTU[ptu$pangenome.category=="include"],]

wilcox_tests = list(wilcox.test(df_4_filter$discrepancy_selected, df_4_filter$discrepancy_others, paired = TRUE),
                    wilcox.test(df_5_filter$discrepancy_selected, df_5_filter$discrepancy_others, paired = TRUE),
                    wilcox.test(df_6_filter$discrepancy_selected, df_6_filter$discrepancy_others, paired = TRUE))
results_df = data.frame(median_discrepancy_targets=c(median(df_4_filter$discrepancy_selected),
                                        median(df_5_filter$discrepancy_selected),
                                        median(df_6_filter$discrepancy_selected)),
           median_discrepancy_others=c(median(df_4_filter$discrepancy_others),
                                        median(df_5_filter$discrepancy_others),
                                        median(df_6_filter$discrepancy_others)),
           wilcoxon_stat=c(wilcox_tests[[1]]$statistic,
                           wilcox_tests[[2]]$statistic,
                           wilcox_tests[[3]]$statistic),
           p_value=c(wilcox_tests[[1]]$p.value,
                     wilcox_tests[[2]]$p.value,
                     wilcox_tests[[3]]$p.value))
kable(results_df, digits = 3, format = "html")

```

### Hard shell vs. accessory for within-range RM targets

```{r}
# Read in results for RM targets
df_4 = read.csv('../results/rmes_discrepancies_targets_k4.csv')
df_4_filter = df_4[df_4$PTU %in% ptu$PTU[ptu$pangenome.category=="include" & df_4$discrepancy_selected!="None"],]
df_5 = read.csv('../results/rmes_discrepancies_targets_k5.csv')
df_5_filter = df_5[df_5$PTU %in% ptu$PTU[ptu$pangenome.category=="include" & df_5$discrepancy_selected!="None"],]
df_6 = read.csv('../results/rmes_discrepancies_targets_k6.csv')
df_6_filter = df_6[df_6$PTU %in% ptu$PTU[ptu$pangenome.category=="include" & df_6$discrepancy_selected!="None"],]

# Convert discrepancies to numeric (presence of None has messed this up)
df_4_filter$discrepancy_selected = as.numeric(df_4_filter$discrepancy_selected)
df_5_filter$discrepancy_selected = as.numeric(df_5_filter$discrepancy_selected)
df_6_filter$discrepancy_selected = as.numeric(df_6_filter$discrepancy_selected)

wilcox_tests = list(wilcox.test(df_4_filter$discrepancy_selected, df_4_filter$discrepancy_others, paired = TRUE),
                    wilcox.test(df_5_filter$discrepancy_selected, df_5_filter$discrepancy_others, paired = TRUE),
                    wilcox.test(df_6_filter$discrepancy_selected, df_6_filter$discrepancy_others, paired = TRUE))
results_df = data.frame(n_PTUs = c(nrow(df_4_filter),
                                   nrow(df_5_filter),
                                   nrow(df_6_filter)),
                        median_n_targets = c(median(df_4_filter$n_selected_kmers),
                                   median(df_5_filter$n_selected_kmers),
                                   median(df_6_filter$n_selected_kmers)),
  median_discrepancy_targets=c(median(df_4_filter$discrepancy_selected),
                                        median(df_5_filter$discrepancy_selected),
                                        median(df_6_filter$discrepancy_selected)),
           median_discrepancy_others=c(median(df_4_filter$discrepancy_others),
                                        median(df_5_filter$discrepancy_others),
                                        median(df_6_filter$discrepancy_others)),
           wilcoxon_stat=c(wilcox_tests[[1]]$statistic,
                           wilcox_tests[[2]]$statistic,
                           wilcox_tests[[3]]$statistic),
           p_value=c(wilcox_tests[[1]]$p.value,
                     wilcox_tests[[2]]$p.value,
                     wilcox_tests[[3]]$p.value))
kable(results_df, digits = 3, format = "html")
```

Repeated this analysis with one gene representative per family:

```{r}
# Read in results for RM targets
df_4 = read.csv('../test_one_per_family_within_range_targets_k4.csv')
df_5 = read.csv('../test_one_per_family_within_range_targets_k5.csv')
df_6 = read.csv('../test_one_per_family_within_range_targets_k6.csv')

# Convert discrepancies to numeric (presence of None has messed this up)
df_4$discrepancy_selected = as.numeric(df_4$discrepancy_selected)
df_4$discrepancy_selected[is.na(df_4$discrepancy_selected)] = 0
df_5$discrepancy_selected = as.numeric(df_5$discrepancy_selected)
df_5$discrepancy_selected[is.na(df_5$discrepancy_selected)] = 0
df_6$discrepancy_selected = as.numeric(df_6$discrepancy_selected)
df_6$discrepancy_selected[is.na(df_6$discrepancy_selected)] = 0

# Median of selected k-mers
df = t(data.frame(c(median(df_4[which(df_4$n_selected_kmers!=0),"discrepancy_selected"]), 
        wilcox.test(df_4[which(df_4$n_selected_kmers!=0),"discrepancy_selected"])$statistic,
        wilcox.test(df_4[which(df_4$n_selected_kmers!=0),"discrepancy_selected"])$p.value),
        c(median(df_5[which(df_5$n_selected_kmers!=0),"discrepancy_selected"]), 
        wilcox.test(df_5[which(df_5$n_selected_kmers!=0),"discrepancy_selected"])$statistic,
        wilcox.test(df_5[which(df_5$n_selected_kmers!=0),"discrepancy_selected"])$p.value),
        c(median(df_6[which(df_6$n_selected_kmers!=0),"discrepancy_selected"]), 
        wilcox.test(df_6[which(df_6$n_selected_kmers!=0),"discrepancy_selected"])$statistic,
        wilcox.test(df_6[which(df_6$n_selected_kmers!=0),"discrepancy_selected"])$p.value)))
colnames(df) = c("median", "statistic", "p.value")
print(df)

```

### All hard shell vs. accessory, one representative per gene family

We subsampled to one representative per gene family and the same number of hard shell and accessory genes, whichever was smaller (PTU-Bac37 was the only one that had more hard shell than accessory genes). 
```{r}
palindromes_k4 = tolower(read.csv('~/Downloads/trieste/k4.txt', header=F)$V1)
palindromes_k6 = tolower(read.csv('~/Downloads/trieste/k6.txt', header=F)$V1)

core_k6 = read.csv('~/Downloads/trieste/one_per_family/all_core_1745k_k6.csv')
accessory_k6 = read.csv('~/Downloads/trieste/one_per_family/all_accessory_1745k_k6.csv')
merge_k6 = merge(core_k6, accessory_k6, by="word")
merge_k6$score.diff = merge_k6$score.x-merge_k6$score.y
merge_k6$count.diff = merge_k6$count.x-merge_k6$count.y

merge_k6$palindrome = ifelse(merge_k6$word %in% palindromes_k6,
                             "palindrome", "other")
median(merge_k6[which(merge_k6$palindrome=="palindrome"),"score.diff"])
wilcox.test(merge_k6[which(merge_k6$palindrome=="palindrome"),"score.diff"])
sum(merge_k6$count.x[which(merge_k6$palindrome=="palindrome")])/sum(merge_k6$count.x) * 1000
sum(merge_k6$count.y[which(merge_k6$palindrome=="palindrome")])/sum(merge_k6$count.y) * 1000

# And for k=4

core_k4 = read.csv('~/Downloads/trieste/one_per_family/all_core_1745k_k4.csv')
accessory_k4 = read.csv('~/Downloads/trieste/one_per_family/all_accessory_1745k_k4.csv')
merge_k4 = merge(core_k4, accessory_k4, by="word")
merge_k4$score.diff = merge_k4$score.x-merge_k4$score.y
merge_k4$count.diff = merge_k4$count.x-merge_k4$count.y

merge_k4$palindrome = ifelse(merge_k4$word %in% palindromes_k4,
                             "palindrome", "other")
median(merge_k4[which(merge_k4$palindrome=="palindrome"),"score.diff"])
wilcox.test(merge_k4[which(merge_k4$palindrome=="palindrome"),"score.diff"])

# Sum palindromes
sum(merge_k4$count.x[which(merge_k4$palindrome=="palindrome")])/sum(merge_k4$count.x) * 1000
sum(merge_k4$count.y[which(merge_k4$palindrome=="palindrome")])/sum(merge_k4$count.y) * 1000

```

## Analysis of leading region

Some preliminary statistics about discarded PTUs (for main text).

```{r}
lr = read.csv('../data/leading_region_081123.tsv', sep='\t')
plasmid.table = read.csv('../data/plasmid_table.tsv', sep='\t')
retained.PTUs.lr = read.csv('../data/leading-region-PTUs.txt', header=F)$V1
lr.identified.PTU.names = sort(unique(lr$PTU))
# Minimum sizes
PTU.min.sizes = sort(sapply(lr.identified.PTU.names, function(x) 
  min(plasmid.table[plasmid.table$PTU==x, "Size"])))
```

We discard those where the minimum size is <20kb: E1, E3, E9, E10, E14, E76. We check the maximum size per PTU:

```{r}
sort(sapply(paste0("PTU-E", c("1", "3", "9", "10", "14", "76")), function(x) 
  max(plasmid.table[plasmid.table$PTU==x, "Size"])))
```

These PTUs are retained for analysing the leading region:

```{r}
leading.region.PTUs = sort(names(PTU.min.sizes[PTU.min.sizes>20000]))
cat(leading.region.PTUs, sep="\n")
```

For convenience for other scripts, I save some files with names of plasmids within each PTU.

```{r}
for (PTU in unique(plasmid.table$PTU)){
  cat(plasmid.table$ID[plasmid.table$PTU==PTU], sep="\n",
      file=paste0('~/Downloads/trieste/', PTU, '_plasmids.txt'))
}

```
### Sliding window analysis

k=6 
```{r}
# Theoretical expectation for 6-bp palindrome density
theoreticalPalindromeExpectation = function(p){# p is GC content
  return ((p**2 - p + 1/2)**3)
}
d = read.csv('../test_output_window5000_step500.csv', header=F, stringsAsFactors = F)
colnames(d) = c("PTU", "plasmid", "x", "window_size", "palindrome_count", "palindrome_density", "GC_n")

d.reduced = d[which(d$PTU %in% leading.region.PTUs), ]

d.reduced$expected_density = theoreticalPalindromeExpectation(d.reduced$GC_n/d.reduced$window_size)

d.reduced$diff_density = d.reduced$palindrome_density-d.reduced$expected_density

# One line per plasmid
d.reduced.median = d.reduced %>% group_by(PTU, x) %>% 
  summarise(median=median(diff_density),
            n=length(diff_density), 
            median.original=median(palindrome_density),
            GC=median(GC_n))

# Maximum counts
d.reduced.median.plot = d.reduced.median %>% 
  group_by(PTU) %>%
  mutate(max_n = max(n, na.rm = TRUE),  # Calculate max 'n' for each 'PTU'
         normalized_n = n / max_n * 100) %>%  # Normalize 'n' by this max value
  ungroup() %>%  # Remove the grouping
  select(-max_n)  # Optionally remove the max_n column if not needed
# Plot the distribution
p.lines.expectation = ggplot(d.reduced.median.plot, aes(x, median*1000))+
  geom_hline(yintercept = 0, linetype='dashed')+
  theme_bw()+
  geom_path(aes(group=PTU, alpha=normalized_n))+
  facet_wrap(~PTU, scales="free_x", nrow=3)+
  ylab("Palindrome depletion per kb (5kb sliding window)")+
  labs(alpha="% of plasmids within PTU")+
  theme(panel.grid=element_blank())+  
  scale_x_continuous(labels = function(x) x / 1000) +
  xlab("Position on plasmid (kb)")+
  theme(legend.position = c(0.8, 0.2))
ggsave(p.lines.expectation, file='../results/6mer-deviation-expectation.pdf',
       width=10, height=6)
# Plot only first 25kb
ggsave(p.lines.expectation+scale_x_continuous(labels = function(x) x / 1000, limits=c(0,50000)), file='../results/6mer-deviation-expectation-0-50-kb.pdf',
       width=10, height=6)

```
```{r}
# Sliding window for GC content
p.lines.GC = ggplot(d.reduced.median.plot, aes(x, GC/5000 * 100))+
  geom_hline(yintercept = 0, linetype='dashed')+
  theme_bw()+
  geom_path(aes(group=PTU, alpha=normalized_n))+
  facet_wrap(~PTU, scales="free_x", nrow=3)+
  ylab("GC content (5kb sliding window)")+
  labs(alpha="% of plasmids within PTU")+
  theme(panel.grid=element_blank())+  
  scale_x_continuous(labels = function(x) x / 1000) +
  xlab("Position on plasmid (kb)")+
  theme(legend.position = c(0.8, 0.2))
ggsave(p.lines.GC, file='../results/GC-content.pdf',
       width=10, height=6)

# And for palindrome density (not expected - actual values)
p.lines.density = ggplot(d.reduced.median.plot, aes(x, median.original*1000))+
  geom_hline(yintercept = 0, linetype='dashed')+
  theme_bw()+
  geom_path(aes(group=PTU, alpha=normalized_n))+
  facet_wrap(~PTU, scales="free_x", nrow=3)+
  ylab("Palindrome density per kb (5kb sliding window)")+
  labs(alpha="% of plasmids within PTU")+
  theme(panel.grid=element_blank())+  
  scale_x_continuous(labels = function(x) x / 1000) +
  xlab("Position on plasmid (kb)")+
  theme(legend.position = c(0.8, 0.2))
ggsave(p.lines.density, file='../results/6mer-palindrome-density.pdf',
       width=10, height=6)

```

### Compare leading region to rest of plasmid

```{r}
# Start from d.reduced, which contains per-plasmid results with 5000bp sliding window and step size of 500bp
# Compare depletion in first 5kb with average for the rest of the plasmid
diff.leading.region.average = d.reduced %>% 
  group_by(plasmid, PTU) %>%
  mutate(diff.count=palindrome_count[x=="0"]-mean(palindrome_count),
         diff.density=diff_density[x=="0"]-mean(diff_density),
         diff.GC=(GC_n[x=="0"]-mean(GC_n))/5000 * 100) %>%
  filter(x=="0")

# Order the PTUs by mean values of depletion for better visual display
mean.values = diff.leading.region.average %>% group_by(PTU) %>%
  summarise(mean=mean(diff.density)*1000)
diff.leading.region.average$PTU = ordered(gsub("PTU-", "", diff.leading.region.average$PTU),
                                          levels=gsub("PTU-", "", mean.values$PTU[order(mean.values$mean)]))
p.leading.region.diff.density = ggplot(diff.leading.region.average, aes(as.factor(PTU), diff.density*1000))+ # depletion per-kb rather than per-bp
  ggbeeswarm::geom_quasirandom(alpha=0.8)+
  theme_bw()+
  theme(panel.grid = element_blank())+
  geom_hline(yintercept = 0)+
  xlab("")+
  theme(axis.text=element_text(colour="black", size=18))+
  ylab("Palindrome depletion in leading region (per kb)")+
  theme(axis.text.x=element_text(angle=45, hjust=1))+
  ylab("")
p.leading.region.diff.density
# Based on this, remove two outliers
p.leading.region.diff.density = p.leading.region.diff.density+
  scale_y_continuous(breaks=seq(-8, 4, 2), limits=c(-8,5))
ggsave(p.leading.region.diff.density, file='../results/6mer-deviation-leading-region-vs-rest-of-plasmid-density.pdf',
       width=5.5, height=6)
```

Now we want to correlate the GC content difference of the first 5kb of the plasmid with the difference in palindrome density

```{r}
diff.PTU.averages = diff.leading.region.average %>% group_by(PTU) %>%
  summarise(mean.GC=mean(diff.GC),
            GC=mean(GC_n[x=="0"]/5000 * 100),
            mean.palindrome=mean(palindrome_density),
            mean.palindrome.diff=mean(diff.density) * 1000,
            n=length(diff.GC))
ggplot(diff.PTU.averages, aes(mean.GC, mean.palindrome.diff))+
  geom_hline(yintercept = 0, linetype='dashed')+
  ggrepel::geom_text_repel(aes(label=PTU))+
    geom_point(aes(size=n), shape=21, colour="black", fill="grey")+
  theme_bw()+
  theme(panel.grid = element_blank())+
  xlab("GC difference between first 5kb\nand plasmid average (%)")+
  ylab("Additional palindrome depletion in leading region (per kb)")

# Calculate actual GC and actual palindrome density
# for a given x-value. I think 5kb is appropriate for leading region to 
# get rid of start effects. 
gc_data_frame = function(position_on_plasmid){
  d.reduced %>% 
  group_by(plasmid, PTU) %>%
  mutate(GC=GC_n[x==position_on_plasmid]/5000 * 100,
         diff.GC=(GC_n[x==position_on_plasmid]-mean(GC_n))/5000 * 100,
         density=palindrome_density[x==position_on_plasmid]*1000,
         density.depletion=diff_density[x==position_on_plasmid]*1000) %>%
  filter(x==position_on_plasmid)
}
d.GC.palindrome = gc_data_frame(5000)
ggplot(d.GC.palindrome, aes(diff.GC, density.depletion))+
  geom_point()

d.GC.palindrome.PTU = d.GC.palindrome %>% group_by(PTU) %>%
  summarise(GC=mean(GC),
            diff.GC=mean(diff.GC),
            density=mean(density),
            density.depletion=mean(density.depletion),
            n=length(window_size))
ggplot(d.GC.palindrome, aes(diff.GC, density.depletion, fill=PTU))+
  geom_point(shape=21, colour="black")+
  facet_wrap(~PTU)
p.gc.correlation = ggplot(d.GC.palindrome, aes(diff.GC, density.depletion))+
  geom_point()+
  theme_bw()+
  xlab("Relative GC content at 5-10kb (%)")+
  ylab("Relative palindrome depletion at 5-10kb (per kb)")+
    xlim(c(-15, 10))+
  ylim(-13, 14)
p.gc.correlation.ptu =  ggplot(d.GC.palindrome.PTU, aes(diff.GC, density.depletion))+
  geom_point(aes(size=n, fill=PTU), shape=21, colour="black")+
  theme_bw()+
  ggrepel::geom_text_repel(aes(label=PTU), size=2, min.segment.length = 0)+
  guides(fill=FALSE)+
    xlab("Relative GC content (%)")+
  ylab("Relative palindrome depletion (per kb)")
pdf('../results/GC-correlation-depletion.pdf', width=8, height=4)
cowplot::plot_grid(p.gc.correlation+ggtitle("(a)"), p.gc.correlation.ptu+ggtitle("(b)"))
dev.off()
cor.test(d.GC.palindrome$diff.GC, d.GC.palindrome$density.depletion, method="spearman")
cor.test(d.GC.palindrome.PTU$diff.GC, d.GC.palindrome.PTU$density.depletion)

summary(lm(density.depletion ~ diff.GC*PTU, data=d.GC.palindrome))

```

### Averaging over all plasmids

Similar to Samuel et al.'s display of anti-defence gene density, average over all conjugative plasmids.

```{r}
overall.median = d.reduced %>% filter(x<50000) %>%
  group_by(x) %>%
  summarise(y=median(diff_density),
            n=length(diff_density))
p.overall.average = ggplot(overall.median, aes(x, y*1000))+
  geom_line()+
  scale_y_continuous(breaks=seq(-10, 0, 2), limits = c(-9, 0))+
  theme_bw()+
  theme(panel.grid = element_blank())+
  geom_hline(yintercept = 0, linetype='dashed')+
  ylab("Palindrome depletion (per kb)")+
    scale_x_continuous(labels = function(x) x / 1000) +
  xlab("Position on plasmid (kb)")+
  theme(axis.text=element_text(colour="black"))
p.overall.average
# Save
ggsave(p.overall.average, file='../results/6mer-deviation-overall.pdf', 
       width=3, height=3)
```
Also do this for GC content.

```{r}
overall.median.GC = d.reduced %>% filter(x<50000) %>%
  group_by(x) %>%
  summarise(y=mean(GC_n),
            n=length(GC_n))
p.overall.average.GC = ggplot(overall.median.GC, aes(x, y/5000 * 100))+
  geom_line()+
  theme_bw()+
  theme(panel.grid = element_blank())+
  geom_hline(yintercept = 50, linetype='dashed')+
  ylab("GC content (per kb)")+
    scale_x_continuous(labels = function(x) x / 1000) +
  xlab("Position on plasmid (kb)")+
  theme(axis.text=element_text(colour="black"))
p.overall.average.GC
# Save
ggsave(p.overall.average.GC, file='../results/GC-mean-overall.pdf', 
       width=5, height=2)
```

```{r}
lagging.df = read.csv('~/Downloads/trieste/leading_lagging/NZ_LT994833.1_lagging_10000_k6.csv', header=T)
leading.df = read.csv('~/Downloads/trieste/leading_lagging/NZ_LT994833.1_leading_10000_k6.csv', header=T)
merged.df = merge(leading.df, lagging.df, by="word")
palindromes = tolower(read.csv('~/Downloads/trieste/k6.txt', header=F)$V1)
ggplot(merged.df[which(merged.df$word %in% palindromes),], aes(score.x, score.y))+
  geom_point()
merged.df.palindromes = merged.df[which(merged.df$word %in% palindromes),]
table(merged.df.palindromes$score.x<merged.df.palindromes$score.y)
kmer_targets = tolower(read.csv('~/Downloads/trieste/PTU-LM_targets.txt', header=F)$V1)
ggplot(merged.df[which(merged.df$word %in% kmer_targets),], aes(score.x, score.y))+
  geom_point()
merged.df.targets = merged.df[which(merged.df$word %in% kmer_targets),]
table(merged.df.targets$score.x<merged.df.targets$score.y)

```

### Palindromes in leading vs. lagging

```{r}
scores_df = read.csv('../test_6mer_leading_lagging_scores.csv', header=T)
scores_average <- scores_df %>%
  rowwise() %>%
  mutate(score_avg = mean(c_across(starts_with("score")))) %>%
  ungroup() %>%
  select(kmer, score_avg)  
# Order by average score
scores_average$kmer = ordered(scores_average$kmer,
                              levels=scores_average$kmer[order(scores_average$score_avg)])
scores.plot = ggplot(scores_average, aes(kmer, score_avg))+geom_point()+
  coord_flip()

# Same for counts
counts_df = read.csv('../test_6mer_leading_lagging_counts.csv', header=T)
counts_average <- counts_df %>%
  rowwise() %>%
  mutate(count_avg = mean(c_across(starts_with("count")))) %>%
  ungroup() %>%
  select(kmer, count_avg)

# Merge together
merge_scores_counts = merge(counts_average, scores_average, by="kmer")
# average score and average count are well-correlated
ggplot(merge_scores_counts, aes(count_avg, score_avg))+geom_point()

# Order by average score
merge_scores_counts$kmer = ordered(merge_scores_counts$kmer,
                              levels=scores_average$kmer[order(scores_average$score_avg)])
scores.plot = ggplot(merge_scores_counts, aes(score_avg, kmer))+
    geom_vline(xintercept = 0)+
  geom_bar(stat="identity")+
  geom_vline(xintercept = median(merge_scores_counts$score_avg), linetype='dashed', colour='red')+
  theme_bw()+
  theme(axis.text=element_text(colour="black", size=12),
        axis.title.x=element_text(size=24),
        axis.text.x=element_text(size=18))+
  theme(axis.text.y=element_text(hjust=0))+
  ylab("")+
  xlab("Mean depletion score\n(leading vs. lagging, 10kb)")
wilcox.test(merge_scores_counts$score_avg)
ggsave(scores.plot, file='../results/6mer-palindromes-scores.pdf',
       height=10, width=10)
```

Test for a particular PTU.

```{r}

results = NULL
for (PTU in leading.region.PTUs){
  for (k in c(5,6)){
    scores_df = read.csv(paste0('../test_target_leading_lagging/', PTU, '_k', k, '_scores.csv'), header=T)
    n_targets = nrow(scores_df)
    scores_average <- scores_df %>%
  rowwise() %>%
  mutate(score_avg = mean(c_across(starts_with("score")))) %>%
  ungroup() %>%
  select(kmer, score_avg)
    counts_df = read.csv(paste0('../test_target_leading_lagging/', PTU, '_k', k, '_counts.csv'), header=T)
    counts_average <- counts_df %>%
  rowwise() %>%
  mutate(count_avg = mean(c_across(starts_with("count")))) %>%
  ungroup() %>%
  select(kmer, count_avg)
    results = rbind(results, c(PTU, k, n_targets, median(scores_average$score_avg), median(counts_average$count_avg)))
  }
}

results = data.frame(results)
colnames(results) = c("PTU", "k", "n.targets", "median.score", "median.count")
results$median.score= as.numeric(results$median.score)
results$median.count = as.numeric(results$median.count)

# Order by median score for k=6
order.k.6.scores = results[which(results$k==6), "PTU"][order(results[which(results$k==6), "median.score"])]
results$PTU = ordered(results$PTU,
                      levels=rev(order.k.6.scores))
# Add number of targets to each

p.depletion.targets = ggplot(results, aes(median.score, as.factor(PTU)))+
  geom_vline(xintercept = 0, size=1)+
  geom_bar(stat="identity", position="dodge")+
  facet_wrap(~k, scales="free_x")+
  ylab("")+
  xlab("Median depletion score of within-range RM targets\n(leading vs. lagging, 10kb)")+
   theme_bw()+
  theme(panel.grid = element_blank())+
  theme(axis.text=element_text(colour="black", size=12),
        axis.title.x=element_text(size=24),
        axis.text.x=element_text(size=18))+
  theme(axis.text.y=element_text(hjust=0))
ggsave(p.depletion.targets, file='../results/figure-depletion-within-range-targets.pdf',
       width=10, height=6)

# Is there a correlation between k=5 and k=6 results?
results.df = results %>% select(k, median.score, PTU) %>%
  pivot_wider(
    names_from = k,  # Use values of 'k' as new column names
    values_from = median.score  # Use 'median.score' for the column values
)
```

### All motifs testing for over-representation

Go through values of k for across PTU averages.

```{r}
# Normalise
PTU.counts = table(plasmid.table$PTU)
n.plasmids = PTU.counts[leading.region.PTUs]
# k= 4
# 6-bp palindromes
palindromes = read.csv('~/Downloads/trieste/k4.txt', header=F)$V1
results.df = read.csv('../test_ptus_k4_scores.csv',header=T)
# Add PTUs as column names
colnames(results.df)[2:ncol(results.df)] = leading.region.PTUs
results.df.norm = results.df %>%
mutate(across(starts_with("PTU-"), ~ .x / n.plasmids[cur_column()], .names = "{.col}"))
# Plot all palindromes
results.df.long = results.df.norm %>%
pivot_longer(cols = starts_with("PTU"),
names_to = "score_type",
values_to = "score_avg")
results.df.norm.average = results.df.long %>% group_by(kmer) %>%
summarise(median=median(score_avg)) %>%
arrange(median)
# Take the top 1% depleted
results.df.norm.average.top.depleted = results.df.norm.average[which(results.df.norm.average$median< quantile(results.df.norm.average$median, probs = c(0.05))),] # not enough for 0.01
results.df.norm.average.top.depleted$palindrome = ifelse(results.df.norm.average.top.depleted$kmer %in% palindromes,
"palindrome",
"non-palindrome")
# Read in all RM targets
all_rm_targets = read.csv('../data/rebase_all_4_6bp_targets_250131.csv', header=T)
all_rm_targets_4bp =all_rm_targets$target[all_rm_targets$length==4]
results.df.norm.average.top.depleted$target = ifelse(results.df.norm.average.top.depleted$kmer %in%
all_rm_targets_4bp,
"target", "non-target")
table(results.df.norm.average.top.depleted$palindrome)
table(results.df.norm.average.top.depleted$target)
# palindromes
chisq.test(matrix(c(2, 11, 13, 230 ), nrow=2))
# Targets
chisq.test(matrix(c(3, 10, 100, 143 ), nrow=2))

### And for k=5
results.df = read.csv('../test_ptus_k5_scores.csv',header=T)
# Add PTUs as column names
colnames(results.df)[2:ncol(results.df)] = leading.region.PTUs
results.df.norm = results.df %>%
mutate(across(starts_with("PTU-"), ~ .x / n.plasmids[cur_column()], .names = "{.col}"))
results.df.long = results.df.norm %>%
pivot_longer(cols = starts_with("PTU"),
names_to = "score_type",
values_to = "score_avg")
results.df.norm.average = results.df.long %>% group_by(kmer) %>%
summarise(median=median(score_avg)) %>%
arrange(median)
# Take the top 1% depleted
results.df.norm.average.top.depleted = results.df.norm.average[which(results.df.norm.average$median< quantile(results.df.norm.average$median, probs = c(0.01))),] 
all_rm_targets_5bp =all_rm_targets$target[all_rm_targets$length==5]
results.df.norm.average.top.depleted$target = ifelse(results.df.norm.average.top.depleted$kmer %in%
all_rm_targets_5bp,
"target", "non-target")
table(results.df.norm.average.top.depleted$target)
chisq.test(matrix(c(5, 11, 277, 731 ), nrow=2))

# k =6
# 6-bp palindromes
palindromes = read.csv('~/Downloads/trieste/k6.txt', header=F)$V1
results.df = read.csv('../test_ptus_k6_scores.csv',header=T)
# Add PTUs as column names
colnames(results.df)[2:ncol(results.df)] = leading.region.PTUs
results.df.norm = results.df %>%
mutate(across(starts_with("PTU-"), ~ .x / n.plasmids[cur_column()], .names = "{.col}"))
# Switch to long format
results.df.long = results.df.norm %>%
pivot_longer(cols = starts_with("PTU"),
names_to = "score_type",
values_to = "score_avg")

results.df.norm.average = results.df.long %>% group_by(kmer) %>%
summarise(median=median(score_avg)) %>%
arrange(median)
# Take the top 1% depleted
results.df.norm.average.top.depleted = results.df.norm.average[which(results.df.norm.average$median< quantile(results.df.norm.average$median, probs = c(0.01))),]
results.df.norm.average.top.depleted$palindrome = ifelse(results.df.norm.average.top.depleted$kmer %in% palindromes,
"palindrome",
"non-palindrome")
# Read in all RM targets
all_rm_targets = read.csv('../data/rebase_all_4_6bp_targets_250131.csv', header=F, sep=' ')
all_rm_targets_6bp =all_rm_targets$target[all_rm_targets$length==6]
results.df.norm.average.top.depleted$target = ifelse(results.df.norm.average.top.depleted$kmer %in%
all_rm_targets_6bp,
"target", "non-target")
table(results.df.norm.average.top.depleted$palindrome)
table(results.df.norm.average.top.depleted$target)
chisq.test(matrix(c(7, 34, 865-7, 4096-865-34 ), nrow=2))
chisq.test(matrix(c(4, 37, 60, 4055 ), nrow=2))

```